\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}

### 2023-09-20 manuscript/manuscriptRW.Rmd
reviewer 2 almost all comments

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbance impacts and changing societal
demands on forest ecosystem services, it is becoming ever more important
to understand how forest structure, composition and function will
change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics were developed over the past decades considering the
impacts of climate [@bugmann2022]. From these studies, it is evident
that we have a reasonably good understanding how to model tree growth
[@vanclay1997; @bugmann1996], and substantial efforts have been
dedicated to improving the representation of tree mortality
[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
less studied process that is often represented rather coarsely in models
[@walck2011; @leishman1992; @price2001]. Which is a clear research gap
in the context of climate-induced forest disturbances and forest
resilience.

Tree regeneration arises from multiple processes such as pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
of a few simple environmental filters, as done in most forest "gap"
models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994].

@@

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree regeneration [@käber2023]. By tree regeneration, we refer to the
passing of trees across a specific diameter threshold ("ingrowth"). We
included models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving this process --- including some that are not incorporated
explicitly in most models, such as deer browsing --- we do not aim for a
detailed statistical evaluation of each model. Instead, we aim to
evaluate the general regeneration patterns and magnitudes simulated by
the models, and benchmark the simulated regeneration niche of multiple
species against empirical data along a wide environmental gradient of
temperature, moisture, and light availability [@grubb1977].

@@

We selected 200 sites from this network as the benchmarking dataset for
the simulation to be representative of the environmental variation
contained in the data. This was achieved by applying k-means clustering
to define 200 clusters of plots from the original set of 869 plots along
the environmental dimensions of temperature, climatic water balance,
soil quality and exposition (i.e., slope and aspect). Each of these
sites featured at least two consecutive measurements. Regeneration
thresholds for these sites differed between diameters of 0 and 10 cm.
For our study, we defined two datasets where one included 165 sites with
a diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds between 7 and 10 cm.

@@

The overarching goal of the experiments was to assess the tree
regeneration as it arises from empirical data against its representation
in a wide range of models of forest dynamics. We define tree
regeneration as the passing of a breast height diameter threshold of 7
or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
group was provided with a detailed protocol [@díaz-yáñez2022] with
instructions how to perform the simulations, providing input variables
on climate and soil conditions, and the list of expected output
variables. Neither were further site information (except for the data
specified below) nor any data on tree regeneration or forest stand
features provided prior to the simulation. That is, the models were run
in "blind flight" mode.

@@

Each of the models reported the regeneration number by sampling 200
times in a 10-year interval for each species and per ha for each of the
200 sites. Multiple samples per site were used to better understand the
simulated variation within each site. This was done using different
strategies, depending on the model, including (1) sampling simulated
data from the same 1 ha plot in the equilibrium over time, (2) sampling
several 1 ha plots from the simulated forest at one specific point in
time (in the equilibrium) or (3) a combination of (1) and (2). This
resulted in 880,000 observations per model that simulated the 11 species
included in the protocol (200 sites, 200 samples per site, 11 species
and two diameter thresholds). For the models that simulated additional
species, their regeneration rates were aggregated as "others".

@@

```{=tex}
\begin{equation}
Reineke = \frac{7}{10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
using two-sided t-tests by considering each of these model features
regarding the observed and simulated regeneration levels and species
diversity, respectively.

@@

Regeneration levels define the structure and composition of future
forests, but it is equally important to correctly identify areas where
regeneration is lacking [@rammer2021]. There are multiple constraints to
the regeneration niche of tree species [@price2001], and therefore the
absence of regeneration is likely to be common [@fortin2007], even over
larger areas such as the one-hectare plots used here. Tree regeneration
data are characterized by zero-inflation. This was clearly evident from
the EuFoRIa dataset, but some of the models did not produce zeros at
all, or featured a very low proportion of zero data. This substantial
difference may be due to the fact that the simulation results were drawn
from equilibrium forests, whereas in reality many of the forest reserves
are recovering from past management activities and have become denser
over the past decades [e.g., @heiri2009], leading to less regeneration
than in an equilibrium situation.

@@

The species diversity of the entire stand was captured better by the
models than the diversity of tree regeneration, and only a few models
overpredicted stand-level diversity. Defining regeneration composition
as being proportional to adult abundance, regardless of productivity,
might be a simple and conservative assumption to safely maintain
relative species abundances [@hanbury-brown2022], although this may be
simplistic. Furthermore, based on the simulation results, there is no
evidence that models with feedback from the canopy captured better the
species diversity of the regeneration than those without feedback, with
the exception of ForCEEPS that significantly improved regeneration
species diversity with the model version that included feedback
(ForCEEPS(f)). The similar performance of models with and without
feedback is likely because the models put more weight on the
regeneration niche arising from abiotic and biotic filters, than from
the habitat niche of the adult trees [@grubb1977].

@@

Competition for light as a strong filter for tree regeneration has been
widely documented [@messier1999; @collet2006; @berdanier2016], but the
models examined here did not reproduce this expectation. However, it is
difficult to measure light availability at large spatial and temporal
scales. We used total stand basal area as a proxy for light availability
[cf. @schmid2021]. However, we were unable to consider light
availability restrictions caused by ground vegetation, which may be an
important filter for some EuForIa stands [@woltjer2008]. We found
pronounced differences in the stand basal area ranges simulated by the
models, but also between models and observations. This made it
impossible to evaluate the regeneration for the extremes of the stand
density ranges in some models. For example, regeneration levels at low
stand densities are relevant to assess how well forests are recovering
e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
the level of the simulated one-hectare samples, average basal area is
typically not very low as long as no larger disturbances occur, which
was explicitly excluded in our simulation protocol. Towards the other
end of the spectrum, i.e., with increasing stand basal area, it would be
reasonable to expect that the regeneration of the different species
would become sparser and drop out entirely under low-light conditions
[@klopcic2012; @zell2019]. However, few models showed this trend, thus
indicating that the relation between regeneration and light availability
is not yet captured correctly in most models. Yet, several of the models
that did feature an increase of regeneration with increasing basal area
include a feedback between seed production of mature trees and
regeneration. Thus, it seems that higher light competition does not
sufficiently compensate for increased seed availability with higher
basal area due to a higher abundance of mature trees.

>>>>

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbance impacts and changing societal
demands on forest ecosystem services, it is becoming ever more important
to understand how forest structure, composition and function will
change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics have been developed over the past decades considering
the impacts of climate [@bugmann2022]. From these studies, it is evident
that we have a reasonably good understanding how to model tree growth
[@vanclay1997; @bugmann1996], and substantial efforts have been
dedicated to improving the representation of tree mortality
[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
less studied process that is often represented rather coarsely in models
[@walck2011; @leishman1992; @price2001]. This presents an important
research gap, particularly in the context of climate-induced forest
disturbances and forest resilience.

Tree regeneration arises from multiple processes including pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring them [as done in classical
forest growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2)
the use of a few simple environmental filters, as done in most forest
"gap" models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994].

@@

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree regeneration [@käber2023]. By tree regeneration, we refer to the
passing of trees across a specific diameter threshold ("ingrowth"). We
included models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving and constraining this process --- including some that are not
incorporated explicitly in most models, such as deer browsing --- we do
not aim for a detailed statistical evaluation of each model. Instead, we
aim to evaluate the general regeneration patterns and magnitudes
simulated by the models, and benchmark the simulated regeneration niche
of multiple species against empirical data along a wide environmental
gradient of temperature, moisture, and light availability [@grubb1977].

@@

We selected 200 sites from this network as the benchmarking dataset for
the simulation to be representative of the environmental variation
contained in the data. This was achieved by applying k-means clustering
to define 200 clusters of plots from the original set of 869 plots along
the environmental dimensions of temperature, climatic water balance,
soil quality, slope and aspect. Each of these sites featured at least
two consecutive measurements. Regeneration minimum size thresholds for
these sites differed between stem diameters of 0 and 10 cm. For our
study, we defined two datasets where one included 165 sites with a
diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds between 7 and 10 cm.

@@

The overarching goal of the simulation experiments was to assess tree
regeneration as it arises from empirical data against its representation
in a wide range of models of forest dynamics. We define tree
regeneration as the passing of a breast height diameter threshold of 7
and 10 cm (synonym: ingrowth). To this end, each modeling group was
provided with a detailed protocol [@díaz-yáñez2022] with instructions
how to perform the simulations, input variables on climate and soil
conditions, and the list of expected output variables. Neither were
further site information (except for the data specified below) nor any
data on tree regeneration or forest stand features provided prior to the
simulation.

@@

Each of the models reported the number of trees crossing the two
diameter thresholds by sampling 200 times in a 10-year interval for each
species and per ha for each of the 200 sites. Multiple samples per site
were used to better understand the simulated variation within each site.
This was done using different strategies, depending on the model,
including (1) sampling simulated data from the same 1 ha plot in the
equilibrium over time, (2) sampling several 1 ha plots from the
simulated forest at one specific point in time (in the equilibrium) or
(3) a combination of (1) and (2). This resulted in 880,000 observations
per model that simulated the 11 species included in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
For the models that simulated additional species, their regeneration
rates were aggregated as "others".

@@

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
using two-sided t-tests by considering each of these model features
regarding the observed and simulated regeneration levels and species
diversity, respectively.

@@

Regeneration levels define the structure and composition of future
forests, but it is equally important to correctly identify areas where
regeneration is lacking [@rammer2021]. There are multiple constraints to
the regeneration niche of tree species [@price2001], and therefore the
absence of regeneration is likely to be common [@fortin2007], even over
larger areas such as the one-hectare plots used here. Tree regeneration
data are characterized by zero-inflation. This was clearly evident from
the EuFoRIa dataset, but some of the models did not produce zeros at
all, or featured a very low proportion of zero data. This substantial
difference may be due to the fact that the simulation results were drawn
from equilibrium forests, whereas in reality many of the forest reserves
are recovering from past management activities and have become denser
over the past decades [e.g., @heiri2009], leading to less regeneration
than in an equilibrium situation. Another possible reason for this
difference may be the exclusion of factors like deer browsing or failure
to accurately measure establishment filters, such as those to site
limitants or competition with herb layers.

@@

The species diversity of the entire stand was captured better by the
models than the diversity of tree regeneration, and only a few models
overpredicted stand-level diversity. Defining regeneration composition
as being proportional to adult abundance, regardless of productivity,
might be a simple and conservative assumption to safely maintain
relative species abundances [@hanbury-brown2022], although this may be
simplistic. Furthermore, based on the simulation results, there is no
evidence that models with feedback from the canopy (in terms of species
composition of regeneration) captured better the species diversity of
the regeneration than those without feedback, with the exception of
ForCEEPS that significantly improved regeneration species diversity with
the model version that included feedback (ForCEEPS(f)). The similar
performance of models with and without feedback is likely because the
models put more weight on the regeneration niche arising from abiotic
and biotic filters, than from the habitat niche of the adult trees
[@grubb1977].

@@

Competition for light as a strong filter for tree regeneration has been
widely documented [@messier1999; @collet2006; @berdanier2016], but the
models examined here did not reproduce this expectation (i.e.,
decreasing regeneration levels with increasing stand basal area).
However, it is difficult to measure light availability at large spatial
and temporal scales. We used total stand basal area as a proxy for light
availability [cf. @schmid2021]. However, we were unable to consider
light availability restrictions caused by ground vegetation, which may
be an important filter for some EuForIa stands [@woltjer2008]. We found
pronounced differences in the stand basal area ranges simulated by the
models, but also between models and observations. This made it
impossible to evaluate the regeneration for the extremes of the stand
density ranges in some models. For example, regeneration levels at low
stand densities are relevant to assess how well forests are recovering
e.g. after gap creation due to disturbance [@seidl2022; @grubb1977]. At
the level of the simulated one-hectare samples, average basal area is
typically not very low as long as no larger disturbances occur, which
was explicitly excluded in our simulation protocol. Towards the other
end of the spectrum, i.e., with increasing stand basal area, it would be
reasonable to expect that the regeneration of the different species
would become sparser and drop out entirely under low-light conditions
[@klopcic2012; @zell2019]. However, few models showed this trend, thus
indicating that the relation between regeneration and light availability
is not yet captured correctly in most models. Yet, several of the models
that did feature an increase of regeneration with increasing basal area
include a feedback between seed production of mature trees and
regeneration. Thus, it seems that higher light competition does not
sufficiently compensate for increased seed availability with higher
basal area due to a higher abundance of mature trees.
